\LARGE \bf The Battle for Filter Supremacy: A Retrospective: The Movie: The Game: The Paper

### 2014-12-12 Report/project_report.tex
Added some MSCKF math

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION} \label{sec:introduction}
\cite{Li:2012:ICRA} \cite{Li:2013:IJRR}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MAPLESS VISION-AIDED INERTIAL NAVIGATION}
The combination of visual and inertial sensors is a powerful tool for autonomous navigation in unknown environments.
Indeed, cameras and inertial measurement units (IMUs) are complementary in several respects.
Since an IMU directly measures accelerations and rotational velocities, these values must be integrated to arrive at a new pose estimate.
However, the noise inherent in the IMU's measurements is included in the integration as well, and consequently the pose estimates can drift unbounded over time.
The addition of a camera is an excellent way to bound this cumulative drift error since the camera's signal-to-noise ratio is highest when the camera is stationary.
On the other hand, cameras are not robust to motion blur induced by large accelerations.
In these cases, the strength of the IMU's signal far exceeds its baseline noise and can be relied upon more heavily in estimating pose changes.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Extended Kalman Filter (EKF)} \label{subsec:ekf}
In the Extended Kalman Filter (EKF) solution, vehicle poses and landmark positions are simultaneously estimated at each time step by augmenting the filter state with landmark positions.
This technique, sometimes referred to as EKF-SLAM, attempts to track pose changes and create a globally consistent map of landmarks by recursively updating the state as new measurements become available.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Sliding Window Filter (SWF)} \label{subsec:slidingwindow}
In contrast to EKF-SLAM, the aim of the SWF is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and landmark positions.
The optimization problem in the Sliding Window Filter (SWF) is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Multi-State Constraint Kalman Filter (MSCKF)} \label{subsec:msckf}
The MSCKF \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of landmarks that are visible across the entire window.
This update step typically occurs when a tracked landmark goes out of view of the camera, but it may also be triggered if the number of vehicle states in the window exceeds some preset threshold.

[MSCKF MATH SOMEWHERE]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{EXPERIMENTS} \label{sec:experiments}
We conducted a comparative study of the MSCKF and SWF algorithms using the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS).
The dataset consists of a rigidly attached stereo camera and IMU observing a set of 20 landmarks as they move along an arbitrary 3D path (Figure \ref{fig:sensorhead}).
The dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the landmark positions (Figure \ref{fig:trajectory_groundtruth}).
Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS} \label{sec:conclusions}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION} \label{sec:introduction}

The combination of visual and inertial sensors is a powerful tool for autonomous navigation in unknown environments.
Indeed, cameras and inertial measurement units (IMUs) are complementary in several respects.
Since an IMU directly measures accelerations and rotational velocities, these values must be integrated to arrive at a new pose estimate.
However, the noise inherent in the IMU's measurements is included in the integration as well, and consequently the pose estimates can drift unbounded over time.
The addition of a camera is an excellent way to bound this cumulative drift error since the camera's signal-to-noise ratio is highest when the camera is stationary.
On the other hand, cameras are not robust to motion blur induced by large accelerations.
In these cases, the strength of the IMU's signal far exceeds its baseline noise and can be relied upon more heavily in estimating pose changes.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{EXTENDED KALMAN FILTER} \label{sec:ekf}
In the Extended Kalman Filter (EKF) solution, vehicle poses and landmark positions are simultaneously estimated at each time step by augmenting the filter state with landmark positions.
This technique, sometimes referred to as EKF-SLAM, attempts to track pose changes and create a globally consistent map of landmarks by recursively updating the state as new measurements become available.
%Landmarks are commonly parametrized as three-dimensional vectors and are initialized using two-view geometry, but alternative parametrization and initialization schemes exist \cite{Sola:2011:IJCV}.
%In this paper we consider only the three-dimensional parametrization and two-view initialization scheme.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SLIDING WINDOW FILTER} \label{sec:slidingwindow}
In contrast to EKF-SLAM, the aim of the Sliding Window Filter (SWF) is not to construct a globally consistent map, but rather to estimate a vehicle's motion by optimizing a sliding window of vehicle poses and landmark positions.
The optimization problem in the Sliding Window Filter (SWF) is typically solved as a non-linear least squares problem using Gauss-Newton optimization or some other algorithm.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MULTI-STATE CONSTRAINT KALMAN FILTER} \label{sec:msckf}
The Multi-State Constraint Kalman Filter (MSCKF) \cite{Mourikis:2007:ICRA} can be thought of as a hybrid of EKF-SLAM and the SWF.
The key idea of the MSCKF is to maintain a sliding window of vehicle poses and to simultaneously update each pose in the window using batch-optimized estimates of landmarks that are visible across the entire window.
This update step typically occurs when a tracked landmark goes out of view of the camera, but it may also be triggered if the number of vehicle states in the window exceeds some preset threshold.

\subsection{State Parametrization}
As described in Section \ref{sec:experiments}, we evaluated the MSCKF using a dataset in which the IMU `measures' gravity-corrected linear velocities rather than raw linear accelerations.
In order to accommodate this alternative state parametrization, the mathematical framework described in this section differs slightly from that described in \cite{Mourikis:2007:ICRA}.

We parametrize the IMU state at time $k$ as the 13-dimensional vector
\begin{align}
    \Vec{x}_{I,k} &= \bbm \Vec{q}_{IG,k} & \Vec{b}_{g,k} & \Vec{b}_{v,k} & \Vec{p}_{G,k}^{IG} \ebm^T
\end{align}
where $\Vec{q}_{IG,k}$ is the unit quaternion representing the rotation from the global frame $\mathcal{F}_G$ to the IMU frame $\mathcal{F}_I$, $\Vec{b}_{g,k}$ is the bias on the gyro measurements, $\Vec{b}_{v,k}$ is the bias on the velocity measurements, and $\Vec{p}_{G,k}^{IG}$ is the vector from the origin of $\mathcal{F}_G$ to the origin of $\mathcal{F}_I$ expressed in $\mathcal{F}_G$ (i.e., the position of the IMU in the global frame).

At time $k$, the state of the MSCKF consists of the current IMU state estimate, and estimates of $N$ past camera poses in which active feature tracks were visible:
\begin{align}
    \hat{\Vec{x}}_k &= \bbm \hat{\Vec{x}}_{I,k} & \hat{\Vec{q}}_{C_1 G} & \hat{\Vec{p}}_G^{C_1 G} & \hdots & \hat{\Vec{q}}_{C_N G} & \hat{\Vec{p}}_G^{C_N G} \ebm ^T
\end{align}.

\subsection{IMU State Estimate Propagation}
The evolution of the mean estimated IMU state $\hat{\Vec{x}}_I$ over time is described by the following differential equations:
\begin{align}
    \dot{\hat{\Vec{q}}}_{IG} &= \frac{1}{2}\boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right)\hat{\Vec{q}}_{IG} \\
    \dot{\hat{\Vec{b}}}_g  &= \Vec{0}_{3\times1} \\
    \dot{\hat{\Vec{b}}}_v &= \Vec{0}_{3\times1} \\
    \dot{\hat{\Vec{p}}}^G_I &= \hat{\Vec{C}}_{I G}^T\left(\hat{\Vec{v}}^G_I - \hat{\Vec{b}}_v \right)
\end{align}
where $\hat{\Vec{C}}_{I G}$ is the rotation matrix corresponding to $\hat{\Vec{q}}_{IG}$ and
\begin{align*}
    \boldsymbol{\Omega}\left(\hat{\boldsymbol{\omega}}\right) &= \bbm -\hat{\boldsymbol{\omega}}^\times & \hat{\boldsymbol{\omega}} \\
                                                                -\hat{\boldsymbol{\omega}}^T & 0
                                                            \ebm 
\end{align*}
with
\begin{align*}
    \hat{\boldsymbol{\omega}} = \boldsymbol{\omega}_m - \hat{\Vec{b}}_g
\end{align*}
and
\begin{align*}
    \hat{\boldsymbol{\omega}}^\times &=   \bbm    0 & -\hat{\omega}_3 & \hat{\omega}_2 \\
                                                \hat{\omega}_3 & 0 & -\hat{\omega}_1 \\
                                                -\hat{\omega}_2 & \hat{\omega}_1 & 0
                                            \ebm .
\end{align*}

In our implementation we use a simple forward Euler integration rather than the fifth-order Runge-Kutta procedure used in \cite{Mourikis:2007:ICRA}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{EXPERIMENTS} \label{sec:experiments}
We conducted a comparative study of the MSCKF and SWF algorithms using the ``Starry Night'' dataset from the University of Toronto Institute for Aerospace Studies (UTIAS).
The dataset consists of a rigidly attached stereo camera and IMU observing a set of 20 landmarks as they move along an arbitrary 3D path (Figure \ref{fig:sensorhead}).
The dataset is well-suited to evaluating SLAM algorithms since accurate ground truth from a Vicon motion capture system is available for both the sensor head motion and the landmark positions (Figure \ref{fig:trajectory_groundtruth}).
Since our algorithms are designed to make use of a monocular camera, we artificially blinded the stereo camera by using measurements from the left camera only.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS} \label{sec:conclusions}
